---
title: "Waffle charts"
subtitle: "indsæt subtitle"
author: "Christian Knudsen"
visible: "false" # sæt til noget andet når oplægget/noten er færdigt
categories:
  - R
  - Visualisation
date: 2018-09-11
toc: true
image: ""
---
A rather popular chart type. Not really my favorite, but I can see how it makes things easier to understand for people who are not used to read and understand charts. The reason for my less than favourable view on waffle charts are probably linked to its overuse in meaningless infographics.

A waffle chart is a grid with squares/cells/icons/whatever, where each cell represents a number of something.

Lets make an example:
```{r}
library(ggplot2)
library(waffle)
```

```{r}
vec <- c(`Category 1 (10)`= 10 , `Category 2 (20)`= 20,
              `Category 3 (25)`= 24, `Category 4 (16)` = 16)

waffle(vec/2, rows=3, size=0.1, 
       colors=c("#c7d4b6", "#a3aabd", "#a0d0de", "#97b5cf"), 
       title="Four different categories of something", 
       xlab="1 square = 2 somethings")

```





One annoyance: waffle wants you to spell colours wrong.

waffle takes a named vector of values, rows sets the number of rows of blocks. Default is 10.

One standard way, is to show a 10x10 grid, where each cell represents 1% of the total:

```{r}
waffle(vec/sum(vec)*100)
```



A bit annoying - waffle rounds the values of the vector, leading to only 98 squares. So you have to make sure your values sum to 100 on your own.

What if you want something else than coloured squares?

A selection of fonts with a lot of icons or "glyphs" can be found in
"awesomefonts". `waffle` includes int, but you will have to install them on
your system on your own.

Run

```{r eval = FALSE}
waffle::install_fa_fonts()
```
Which, on my current system returns, amongst other stuff:

`The TTF font files for Font Awesome 5 fonts are in:

/home/christian/R/aarch64-unknown-linux-gnu-library/4.4/waffle/fonts

Please navigate to that directory and install them on your system.`

I navigate there (and get the fonts found there):

```{eval = FALSE}
cd "/home/christian/R/aarch64-unknown-linux-gnu-library/4.4/waffle/fonts"
ls
```

We should get something like this, depending on specific version:

```{eval = FALSE}
Font Awesome 5 Brands-Regular-400.ttf
Font Awesome 5 Free-Regular-400.ttf
Font Awesome 5 Free-Solid-900.ttf
```

Copy those files to the folder containing your fonts. Begin by making the folder:

```{r}
sudo mkdir -p /usr/share/fonts/truetype/fontawesome5
sudo cp "/home/christian/R/aarch64-unknown-linux-gnu-library/4.4/waffle/fonts/"*.ttf /usr/share/fonts/truetype/fontawesome5/
sudo chmod 644 /usr/share/fonts/truetype/fontawesome5/*.ttf
sudo fc-cache -f -v
```

We also adjust permissions, and update the font info cache

Back in R, we load the library `extrafont`, and import the new fonts to R:
```{r eval = FALSE}
library(extrafont)
font_import(paths = "/usr/share/fonts/truetype/fontawesome5", prompt = FALSE)
loadfonts()

```

Just to make sure it actually worked:
```{r eval = FALSE}
"Font Awesome 5 Free Solid" %in% fonts()
```






I needed to place the “fontawesome.ttf” file in the “/usr/share/fonts/truetype/fontawesome” directory.

Then, running R as superuser on the commandline, I imported the extrafont library, and then ran “font_import()”.

But then it worked!

There is now a long list of 593 different icons, that can be used. If you want a list, just run fa_list().

And now, we can make a waffle chart with the glyph of our choice.


```{r eval = FALSE}
font_add("fa", "/usr/share/fonts/truetype/font-awesome/fontawesome-webfont.ttf")
waffle(vec/2, rows=4, use_glyph = "wifi")

install.packages("showtext")
library(showtext)
font_add("fa", "/usr/share/fonts/truetype/fontawesome/fontawesome-webfont.ttf")
showtext_auto(FALSE)
```


## Warning: Removed 1 rows containing missing values (geom_text).


We can change the colours:

library(RColorBrewer)
waffle(vec/2, rows=4, use_glyph = "wifi", colors=brewer.pal(4,"Set1"))

## Warning: Removed 1 rows containing missing values (geom_text).


Adjust the size

waffle(vec/2, rows=4, use_glyph = "wifi", colors=brewer.pal(6,"Set1"), glyph_size=5)

## Warning: Removed 1 rows containing missing values (geom_text).



But that does not look very good.

waffle is based on ggplot, so we have access to the full range of functionality. But not all of them are going to look good in this context:

waffle(vec/2, rows=4, use_glyph = "wifi", colors=brewer.pal(4,"Set1")) +
  geom_label(label="42", size = 3)

## Warning: Removed 1 rows containing missing values (geom_text).


If we install a font that supports it, we even get access to the large number of UTF-8 glyphs. Here is a favorite of mine:

waffle(vec/2, rows=4, colors=brewer.pal(4,"Set1")) +
  geom_label(label=sprintf("\U1F427"), size = 8)


Which of course requires you to have a font on your computer that supports penguins.

Here is one:
http://users.teilar.gr/~g1951d/Symbola.zip



install_fa_fonts()
The TTF font files for Font Awesome 5 fonts are in:

/home/christian/R/aarch64-unknown-linux-gnu-library/4.4/waffle/fonts

Please navigate to that directory and install them on your system.

så gør vi ddet : 
ls "/home/christian/R/aarch64-unknown-linux-gnu-library/4.4/waffle/fonts"

og får noget i denne retning (en af dem mangler hos mig)=
Font Awesome 5 Brands-Regular-400.ttf
Font Awesome 5 Free-Regular-400.ttf
Font Awesome 5 Free-Solid-900.ttf

kopier til fontmappe
sudo mkdir -p /usr/share/fonts/truetype/fontawesome5
sudo cp "/home/christian/R/aarch64-unknown-linux-gnu-library/4.4/waffle/fonts/"*.ttf /usr/share/fonts/truetype/fontawesome5/
sudo chmod 644 /usr/share/fonts/truetype/fontawesome5/*.ttf
sudo fc-cache -f -v

importer i R med extrafont
library(extrafont)
font_import(paths = "/usr/share/fonts/truetype/fontawesome5", prompt = FALSE)
loadfonts()

tjek at den er der
"Font Awesome 5 Free Solid" %in% fonts()

